\documentclass[a4paper,twoside,notitlepage,openany]{tufte-handout}
\usepackage[utf8]{inputenc}

%\geometry{showframe}% for debugging purposes -- displays the margins

% Set up index
\usepackage{makeidx}
\makeindex

% The following package makes prettier tables. We're all about the
% bling!
\usepackage{booktabs}

% The units package provides nice, non-stacked fractions and better
% spacing for units.
\usepackage{units}

% Bold and index keywords
\newcommand{\key}[2][\empty]{\ifthenelse{\equal{#1}{\empty}}
  {\label{key:#2}\textbf{\index{#2}#2}}
  {\label{key:#1}\textbf{\index{#1}#2}}}

% Ordinals: This is the 2\nd of December. Or the 4\nth of July. Or
% 3\rd place in a race.
\usepackage{xspace}
\newcommand\st{\textsuperscript{st}\xspace}
\newcommand\nd{\textsuperscript{nd}\xspace}
\newcommand\rd{\textsuperscript{rd}\xspace}
\renewcommand\th{\textsuperscript{th}\xspace} % override þ

% Clickable hyperlinks
\usepackage{hyperref}
\hypersetup{colorlinks}

\renewcommand{\labelitemi}{$-$} % use dash for items

\title{Halberds \& Helmets: Clerics}
\author{Alex Schroeder}
% \date{}
\begin{document}

\maketitle% this prints the handout title, author, and date

\begin{abstract}
\noindent Clerics are generally not available in my
games.\marginnote[-1em]{There's a long discussion in German
  \href{https://alexschroeder.ch/wiki/2016-01-17_Keine_Kleriker}{on my
    blog}.}
\end{abstract}

\noindent\hyperref[sec:cleric]{Cleric}: you can fight (but not as well
as a fighter) and you can work miracles granted to you by the gods
(but your spells are not as powerful as a wizard’s spells)
\marginnote[-3em]{\emph{What heresy infuriates you the most? When did
    you first hear the divine messengers?}}

\subsection{\index{Clerics}Clerics}
\label{sec:cleric}

\marginnote{
\begin{tabular}{cc}
Level & Experience Points \\
1 & 0\\
2 & 1'565\\
3 & 3'125\\
4 & 6'251\\
5 & 12'501\\
6 & 25'001\\
7 & 50'001\\
8 & 100'001\\
9 & 200'001\\
10 & 300'001\\
\end{tabular}
} You can fight (but not as well as a fighter) and you can work
\hyperref[sec:miracles]{miracles} (but your miracles are not as
powerfull as a magic user's spells).

\smallskip
\begin{tabular}{lcccccccccc}
Armor Class & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
Level 1--3  & 19 & 18 & 17 & 16 & 15 & 14 & 13 & 12 & 11 & 10 \\
Level 4--5  & 18 & 17 & 16 & 15 & 14 & 13 & 12 & 11 & 10 & 9 \\
Level 6--8  & 17 & 16 & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
Level 9--10 & 16 & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 & 7 
\end{tabular}

\medskip\noindent 

\medskip
\begin{tabular}{lcccccccccc}
Saving      & Dragon Breath & Paralysis    & Death  & Rays  & Spells \\
Throws      & Explosions   & Petrification & Poison & Wands & Miracles \\
Level 1--4  & 16 & 14 & 11 & 12 & 15 \\
Level 5--8  & 14 & 12 &  9 & 10 & 12 \\
Level 9--10 & 12 & 10 &  7 &  8 &  9 \\
\end{tabular}

\medskip\noindent You may only use simple, bludgeoning wepons -- or the
weapon ascribed to your god. This means: no magic swords!
\marginnote[-1em]{Some clerics favor the weapon ascribed to their god.
  This is optional. Here are some examples:

\medskip
\begin{tabular}{ll}
\index{Orcus}Orcus & mace \\
\index{Ishtar}Ishtar & sickle sword (kopesh) \\
\index{Nergal}Nergal & battle axe \\
\index{Freya}Freya & spear \\
\index{Marduk}Marduk & club \\
\index{Mitra}Mitra & hand axe \\
\index{Set}Set & dagger \\
\index{Hekate}Hekate & whip 
\end{tabular}
}

\newthought{You can turn undead!}\index{Turn Undead}\index{Undead} You
must raise your holy symbol so they may see it and loudly invoke your
god. Each round, 2d6 hit dice worth of undead will be turned or
destroyed. Your success depends on your level and luck.

The following table shows, what number you need to roll using 2d6 in
order to turn undead. ``D'' means the undead will be ``destroyed''
instead of turned.

\medskip
\begin{tabular}{lcccccccccc}
Cleric Level & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
Skeletons & 7+ & 5+ & 3+ & D & D & D & D & D & D & D \\
Zombies   & 9+ & 7+ & 5+ & 3+ & D & D & D & D & D & D \\
Ghouls    & 11+ & 9+ & 7+ & 5+ & 3+ & D & D & D & D & D \\
Draugr    & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D & D & D & D \\
Wraiths   & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D & D & D \\
Mummies   & -- & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D & D \\
Spectres  & -- & -- & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D \\
Vampires  & -- & -- & -- & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ \\
\end{tabular}

\newpage



\section{\index{Miracles}Miracles}
\label{sec:miracles}

You can beg the gods for a \textbf{miracle} by raising your hands and
praying or singing. The following table shows how your miracle working
improves as you gain levels. You need no spell book. The following
table shows you many miracles you can work per day, depending on your
level.\marginnote{As you can see, miracles are very similar to spells
  for magic users and elves. The key difference is that a cleric's
  \key[Repertoire!Clerics]{repertoire} is determined by the god they
  serve.}

\medskip
\begin{tabular}{cccccc}
Level & 1\st circle & 2\nd circle & 3\rd circle & 4\th circle & 5\th circle \\
1 & 1 & - & - & - & - \\
2 & 2 & - & - & - & - \\
3 & 2 & 1 & - & - & - \\
4 & 3 & 2 & - & - & - \\
5 & 3 & 2 & 1 & - & - \\
6 & 3 & 3 & 2 & - & - \\
7 & 4 & 3 & 2 & 1 & - \\
8 & 4 & 3 & 3 & 2 & - \\
9 & 4 & 4 & 3 & 2 & 1 \\
10 & 5 & 4 & 3 & 3 & 2 \\
\end{tabular}
\marginnote{If you \textbf{anger the gods}, they may deprive you of
  miracles. \emph{Pay attention to dreams and visions} and promise to
  do better in order to regain your full power.}

\medskip Your \textbf{reputation} with the gods determines the highest
circle of miracles granted to you. As you begin the game with no
reputation, you will not be able to work any miracles on your first
adventure!

Miracles \textbf{from the 4\th circle and up} are granted to you
\emph{in person} by agents of your god -- a valkyrie, an angel, a
devil, and so on.

\newthought{Typical miracles} of the 1\st circle would be:

\textbf{Detect Evil} will indicate which visible creatures within
\emph{\unit[60]{ft}} are wishing you ill for
\emph{\unit[1]{h}}.\marginnote{A trap is no creature and doesn't
  ``wish you ill.''}

\textbf{Lay on Hands} heals 1d6+1 damage \emph{or} removes paralysis.
The opposite of that is \textbf{Pain} which causes 1d6+1 damage. In
either case, the target needs to be touched. This does not require a
roll to hit.

\textbf{Detect Magic } makes all enchanted items, creatures and
locations within \emph{\unit[60]{ft}} shine with a pale blue light for
\emph{\unit[20]{min}}.\marginnote[-1em]{The pale blue light is visible
  to all.}

\end{document}
