\documentclass[a4paper,twoside,notitlepage,openany]{tufte-handout}
\usepackage[utf8]{inputenc}

%\geometry{showframe}% for debugging purposes -- displays the margins

% Set up index
\usepackage{makeidx}
\makeindex

% The following package makes prettier tables. We're all about the
% bling!
\usepackage{booktabs}

% The units package provides nice, non-stacked fractions and better
% spacing for units.
\usepackage{units}

% Bold and index keywords
\newcommand{\key}[2][\empty]{\ifthenelse{\equal{#1}{\empty}}
  {\label{key:#2}\textbf{\index{#2}#2}}
  {\label{key:#1}\textbf{\index{#1}#2}}}

% Ordinals: This is the 2\nd of December. Or the 4\nth of July. Or
% 3\rd place in a race.
\usepackage{xspace}
\newcommand\st{\textsuperscript{st}\xspace}
\newcommand\nd{\textsuperscript{nd}\xspace}
\newcommand\rd{\textsuperscript{rd}\xspace}
\renewcommand\th{\textsuperscript{th}\xspace} % override þ

% Clickable hyperlinks
\usepackage{hyperref}
\hypersetup{colorlinks}

\renewcommand{\labelitemi}{$-$} % use dash for items

\title{Halberds \& Helmets: Clerics}
\author{Alex Schroeder}
% \date{}
\begin{document}

\maketitle% this prints the handout title, author, and date

\begin{abstract}
\noindent Clerics are generally not available in my
games.\marginnote[-1em]{There's a long discussion in German
  \href{https://alexschroeder.ch/wiki/2016-01-17_Keine_Kleriker}{on my
    blog}.}
\end{abstract}

\noindent\hyperref[sec:cleric]{Cleric}: you can fight (but not as well
as a fighter) and you can work miracles granted to you by the gods
(but your spells are not as powerful as a wizard’s spells)
\marginnote[-3em]{\emph{What heresy infuriates you the most? When did
    you first hear the divine messengers?}}

\subsection{\index{Clerics}Clerics}
\label{sec:cleric}

\marginnote{
\begin{tabular}{cc}
Level & Experience Points \\
1 & 0\\
2 & 1'565\\
3 & 3'125\\
4 & 6'251\\
5 & 12'501\\
6 & 25'001\\
7 & 50'001\\
8 & 100'001\\
9 & 200'001\\
10 & 300'001\\
\end{tabular}
} You can fight (but not as well as a fighter) and you can work
\hyperref[sec:miracles]{miracles} (but your miracles are not as
powerfull as a magic user's spells).

\smallskip
\begin{tabular}{lcccccccccc}
Armor Class & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
Level 1--3  & 19 & 18 & 17 & 16 & 15 & 14 & 13 & 12 & 11 & 10 \\
Level 4--5  & 18 & 17 & 16 & 15 & 14 & 13 & 12 & 11 & 10 & 9 \\
Level 6--8  & 17 & 16 & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 \\
Level 9--10 & 16 & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 & 7 
\end{tabular}

\medskip\noindent 

\medskip
\begin{tabular}{lcccccccccc}
Saving      & Dragon Breath & Paralysis    & Death  & Rays  & Spells \\
Throws      & Explosions   & Petrification & Poison & Wands & Miracles \\
Level 1--4  & 16 & 14 & 11 & 12 & 15 \\
Level 5--8  & 14 & 12 &  9 & 10 & 12 \\
Level 9--10 & 12 & 10 &  7 &  8 &  9 \\
\end{tabular}

\medskip\noindent You may only use simple, bludgeoning wepons -- or the
weapon ascribed to your god. This means: no magic swords!
\marginnote[-1em]{Some clerics favor the weapon ascribed to their god.
  This is optional. Here are some examples:

\medskip
\begin{tabular}{ll}
\index{Orcus}Orcus & mace \\
\index{Ishtar}Ishtar & sickle sword (kopesh) \\
\index{Nergal}Nergal & battle axe \\
\index{Freya}Freya & spear \\
\index{Marduk}Marduk & club \\
\index{Mitra}Mitra & hand axe \\
\index{Set}Set & dagger \\
\index{Hekate}Hekate & whip 
\end{tabular}
}

\newthought{You can turn undead!}\index{Turn Undead}\index{Undead} You
must raise your holy symbol so they may see it and loudly invoke your
god. Each round, 2d6 hit dice worth of undead will be turned or
destroyed. Your success depends on your level and luck.

The following table shows, what number you need to roll using 2d6 in
order to turn undead. ``D'' means the undead will be ``destroyed''
instead of turned.

\medskip
\begin{tabular}{lcccccccccc}
Cleric Level & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\
Skeletons & 7+ & 5+ & 3+ & D & D & D & D & D & D & D \\
Zombies   & 9+ & 7+ & 5+ & 3+ & D & D & D & D & D & D \\
Ghouls    & 11+ & 9+ & 7+ & 5+ & 3+ & D & D & D & D & D \\
Draugr    & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D & D & D & D \\
Wraiths   & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D & D & D \\
Mummies   & -- & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D & D \\
Spectres  & -- & -- & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ & D \\
Vampires  & -- & -- & -- & -- & -- & 11+ & 9+ & 7+ & 5+ & 3+ \\
\end{tabular}

\newpage



\section{\index{Miracles}Miracles}
\label{sec:miracles}

You can beg the gods for a \textbf{miracle} by raising your hands and
praying or singing. The following table shows how your miracle working
improves as you gain levels. You need no spell book. The following
table shows you many miracles you can work per day, depending on your
level.\marginnote{As you can see, miracles are very similar to spells
  for magic users and elves. The key difference is that a cleric's
  \key[Repertoire!Clerics]{repertoire} is determined by the god they
  serve.}

\medskip
\begin{tabular}{cccccc}
Level & 1\st circle & 2\nd circle & 3\rd circle & 4\th circle & 5\th circle \\
1 & 1 & - & - & - & - \\
2 & 2 & - & - & - & - \\
3 & 2 & 1 & - & - & - \\
4 & 3 & 2 & - & - & - \\
5 & 3 & 2 & 1 & - & - \\
6 & 3 & 3 & 2 & - & - \\
7 & 4 & 3 & 2 & 1 & - \\
8 & 4 & 3 & 3 & 2 & - \\
9 & 4 & 4 & 3 & 2 & 1 \\
10 & 5 & 4 & 3 & 3 & 2 \\
\end{tabular}
\marginnote{If you \textbf{anger the gods}, they may deprive you of
  miracles. \emph{Pay attention to dreams and visions} and promise to
  do better in order to regain your full power.}

\medskip Your \textbf{reputation} with the gods determines the highest
circle of miracles granted to you. As you begin the game with no
reputation, you will not be able to work any miracles on your first
adventure!

Miracles \textbf{from the 4\th circle and up} are granted to you
\emph{in person} by agents of your god -- a valkyrie, an angel, a
devil, and so on.

\newthought{Typical miracles} of the 1\st circle would be:

\textbf{Detect Evil} will indicate which visible creatures within
\emph{\unit[60]{ft}} are wishing you ill for
\emph{\unit[1]{h}}.\marginnote{A trap is no creature and doesn't
  ``wish you ill.''}

\textbf{Lay on Hands} heals 1d6+1 damage \emph{or} removes paralysis.
The opposite of that is \textbf{Pain} which causes 1d6+1 damage. In
either case, the target needs to be touched. This does not require a
roll to hit.

\textbf{Detect Magic } makes all enchanted items, creatures and
locations within \emph{\unit[60]{ft}} shine with a pale blue light for
\emph{\unit[20]{min}}.\marginnote[-1em]{The pale blue light is visible
  to all.}


\clearpage

\section*{License: Open Game License Version 1.0a}

\begin{fullwidth}

\fontsize{6.5pt}{7pt}\selectfont

\textbf{Designation of Open Game Content}: The entire document is open
game content. \textbf{Designation of Product Identity}: The names of
contributors, the name of the website (``Campaign Wiki'') and the name
of this site (``Halberds and Helmets'' as well as ``Hellebarden und
Helme'') are product identity. The following text is the property of
Wizards of the Coast, Inc. and is Copyright 2000 Wizards of the Coast,
Inc (``Wizards''). All Rights Reserved.

\begin{enumerate}

\item Definitions: (a) ``Contributors'' means the copyright and/or
  trademark owners who have contributed Open Game Content; (b)
  ``Derivative Material'' means copyrighted material including
  derivative works and translations (including into other computer
  languages), potation, modification, correction, addition, extension,
  upgrade, improvement, compilation, abridgment or other form in which
  an existing work may be recast, transformed or adapted; (c)
  ``Distribute'' means to reproduce, license, rent, lease, sell,
  broadcast, publicly display, transmit or otherwise distribute; (d)
  ``Open Game Content'' means the game mechanic and includes the
  methods, procedures, processes and routines to the extent such
  content does not embody the Product Identity and is an enhancement
  over the prior art and any additional content clearly identified as
  Open Game Content by the Contributor, and means any work covered by
  this License, including translations and derivative works under
  copyright law, but specifically excludes Product Identity. (e)
  ``Product Identity'' means product and product line names, logos and
  identifying marks including trade dress; artifacts; creatures
  characters; stories, storylines, plots, thematic elements, dialogue,
  incidents, language, artwork, symbols, designs, depictions,
  likenesses, formats, poses, concepts, themes and graphic,
  photographic and other visual or audio representations; names and
  descriptions of characters, spells, enchantments, personalities,
  teams, personas, likenesses and special abilities; places,
  locations, environments, creatures, equipment, magical or
  supernatural abilities or effects, logos, symbols, or graphic
  designs; and any other trademark or registered trademark clearly
  identified as Product identity by the owner of the Product Identity,
  and which specifically excludes the Open Game Content; (f)
  ``Trademark'' means the logos, names, mark, sign, motto, designs
  that are used by a Contributor to identify itself or its products or
  the associated products contributed to the Open Game License by the
  Contributor (g) ``Use'', ``Used'' or ``Using'' means to use,
  Distribute, copy, edit, format, modify, translate and otherwise
  create Derivative Material of Open Game Content. (h) ``You'' or
  ``Your'' means the licensee in terms of this agreement.

\item The License: This License applies to any Open Game Content that
  contains a notice indicating that the Open Game Content may only be
  Used under and in terms of this License. You must affix such a
  notice to any Open Game Content that you Use. No terms may be added
  to or subtracted from this License except as described by the
  License itself. No other terms or conditions may be applied to any
  Open Game Content distributed using this License.

\item Offer and Acceptance: By Using the Open Game Content You
  indicate Your acceptance of the terms of this License.

\item Grant and Consideration: In consideration for agreeing to use
  this License, the Contributors grant You a perpetual, worldwide,
  royalty-free, non-exclusive license with the exact terms of this
  License to Use, the Open Game Content.

\item Representation of Authority to Contribute: If You are
  contributing original material as Open Game Content, You represent
  that Your Contributions are Your original creation and/or You have
  sufficient rights to grant the rights conveyed by this License.

\item Notice of License Copyright: You must update the COPYRIGHT
  NOTICE portion of this License to include the exact text of the
  COPYRIGHT NOTICE of any Open Game Content You are copying, modifying
  or distributing, and You must add the title, the copyright date,
  and the copyright holder's name to the COPYRIGHT NOTICE of any
  original Open Game Content you Distribute.

\item Use of Product Identity: You agree not to Use any Product
  Identity, including as an indication as to compatibility, except as
  expressly licensed in another, independent Agreement with the owner
  of each element of that Product Identity. You agree not to indicate
  compatibility or co-adaptability with any Trademark or Registered
  Trademark in conjunction with a work containing Open Game Content
  except as expressly licensed in another, independent Agreement with
  the owner of such Trademark or Registered Trademark. The use of any
  Product Identity in Open Game Content does not constitute a
  challenge to the ownership of that Product Identity. The owner of
  any Product Identity used in Open Game Content shall retain all
  rights, title and interest in and to that Product Identity.

\item Identification: If you distribute Open Game Content You must
  clearly indicate which portions of the work that you are
  distributing are Open Game Content.

\item Updating the License: Wizards or its designated Agents may
  publish updated versions of this License. You may use any authorized
  version of this License to copy, modify and distribute any Open Game
  Content originally distributed under any version of this License.

\item Copy of this License: You MUST include a copy of this License
  with every copy of the Open Game Content You Distribute.

\item Use of Contributor Credits: You may not market or advertise the
  Open Game Content using the name of any Contributor unless You have
  written permission from the Contributor to do so.

\item Inability to Comply: If it is impossible for You to comply with
  any of the terms of this License with respect to some or all of the
  Open Game Content due to statute, judicial order, or governmental
  regulation then You may not Use any Open Game Material so affected.

\item Termination: This License will terminate automatically if You
  fail to comply with all terms herein and fail to cure such breach
  within 30 days of becoming aware of the breach. All sublicenses
  shall survive the termination of this License.

\item Reformation: If any provision of this License is held to be
  unenforceable, such provision shall be reformed only to the extent
  necessary to make it enforceable.

\item COPYRIGHT NOTICE
\newline
Open Game License v 1.0 © 2000, Wizards of the Coast, Inc.
\newline
System Reference Document Copyright 2000, Wizards of the Coast, Inc.;
Authors Jonathon Tweet, Monte Cook, Skip Williams, based on original
material by E. Gary Gygax and Dave Arneson.
\newline
System Reference Document Copyright 2000-2003, Wizards of the Coast,
Inc.; Authors Jonathan Tweet, Monte Cook, Skip Williams, Rich Baker,
Andy Collins, David Noonan, Rich Redman, Bruce R. Cordell, John D.
Rateliff, Thomas Reid, James Wyatt, based on original material by E.
Gary Gygax and Dave Arneson.
\newline
Modern System Reference Document Copyright 2002-2004, Wizards of the
Coast, Inc.; Authors Bill Slavicsek, Jeff Grubb, Rich Redman, Charles
Ryan, Eric Cagle, David Noonan, Stan!, Christopher Perkins, Rodney
Thompson, and JD Wiker, based on material by Jonathan Tweet, Monte
Cook, Skip Williams, Richard Baker, Peter Adkison, Bruce R. Cordell,
John Tynes, Andy Collins, and JD Wiker.
\newline
Castles \& Crusades: Players Handbook, Copyright 2004, Troll Lord
Games; Authors Davis Chenault and Mac Golden.
\newline
Cave Cricket from the Tome of Horrors, copyright 2002, Necromancer
Games, Inc.; Authors Scott Greene and Clark Peterson, based on
original material by Gary Gygax.
\newline
Crab, Monstrous from the Tome of Horrors, copyright 2002, Necromancer
Games, Inc.; Author Scott Greene, based on original material by Gary
Gygax.
\newline
Fly, Giant from the Tome of Horrors, copyright 2002, Necromancer
Games, Inc.; Author Scott Greene, based on original material by Gary
Gygax.
\newline
Golem, Wood from the Tome of Horrors, copyright 2002, Necromancer
Games, Inc.; Authors Scott Greene and Patrick Lawinger.
\newline
Kamadan from the Tome of Horrors, copyright 2002, Necromancer Games,
Inc.; Author Scott Greene, based on original material by Nick Louth.
\newline
Rot Grub from the Tome of Horrors, Copyright 2002, Necromancer Games,
Inc.; Authors Scott Greene and Clark Peterson, based on original
material by Gary Gygax
\newline
\href{http://www.goblinoidgames.com/labyrinthlord.html}{Labyrinth
  Lord}\textsuperscript{\texttrademark} Copyright 2007-2009, Daniel
Proctor. Author Daniel Proctor.
\newline
\href{https://github.com/kensanata/halberdsnhelmets/tree/master/Hellebarden\%20und\%20Helme}{Hellebarden
  und Helme} Copyright 2012--2014, Alex Schroeder. Author Alex
Schroeder, Samuel Dennler.

\end{enumerate}

\noindent For purposes of this license, nothing is considered to be
Product Identity in addition to anything covered in section 1, above.

\end{fullwidth}


\end{document}

% Local Variables:
% ispell-local-dictionary: "swiss"
% End:
